\pkg{Rgbp}: Bayesian Hierarchical Modeling and Frequentist Method Check

### 2013-05-29 jss/article.tex
fix spelling

Considering the type of data, it makes sense to assume the number of deaths in each hospital has independent Poisson distribution given the unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where $n_{j}$ can be interpreted as an exposure (not necessarily an interger) and $\lambda_{j}$ as a true death rate per exposure. This data set is also included in the package and can be called by `\code{R> data(hospital)}' on \proglang{R}.

@@

\section[Multilevel Structure]{Multilevel Structure}
Our multilevel model, also called a conditionally independent hierarchical model (Kass and Steffey, 1989), is a very powerful tool for exploring the hierarchical sturucture in data. For example, we can think about a district-level hierarchy (bigger population) for 8 schools, the state-level hierarchy for 31 hospitals, and the position-level hierarchy for 18 baseball players. \code{gbp}, one of functions in \pkg{Rgbp}, fits such a hierarchical model whose first-level hierarchy has a distribution of observed data and second-level (bigger population hierarchy) has a conjugate prior distribution on the first-level parameter. Users can determine one of three types of multilevel models, such as Normal-Normal, Poisson-Gamma, and Binomial-Beta, based on their data sets. 
\\

@@

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only data. Taking the Normal model as an example,  $E(\mu_{j}\vert \textbf{y})=E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Adam's law and $Var(\mu_{j}\vert \textbf{y})=E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Eve's law. Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are functions of shrinkage (see (3)) and Adam's and Eve's laws are taking conditional expectation and variance over them given data, $\textbf{y}$. Since we have already estimated $E(B_{j}\vert\textbf{y})$ and $Var(B_{j}\vert\textbf{y})$ via ADM, fianlly we can estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$.

@@

Once we are sure that the hierarchical modeling can be appropriate for our data, the following quetsion will be about the validity of interval estimates, the final product of this multilevel modeling. ``Does the 95\% (can be specified differently) confidence interval obtained via this Bayesian model-fitting process achieve 95\% confidence level  for any true parameter values?'' Our answer is ``yes'' and \pkg{Rgbp} has a function to check this point. From now on, all the explanations will be based on the Binomial model because the idea is the same and can be easily applied to the other two models.

@@

\subsection{Rao-Blackwellized Unbiased estimates of Coverage Probabilities}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking conditional expectation given a sufficient statistic. Based on this idea, we define $E(I_{ij}\vert z_{ij}, r, \beta$), where $r$ and $\beta$ were given at first (see 5.1) and $z_{ij}$ is a sufficient statistic. This expectation is the same as Pr$(\hat{p}_{ij, low}\le p_{ij} \le\hat{p}_{ij, upp}\vert z_{ij}, r, \beta)$, where ($\hat{p}_{ij, low}$, $\hat{p}_{ij, upp}$) is $j$-th player's interval estimates on the $i$-th data set. We can calculate this probability  because we know the distribution of $p_{ij} \vert z_{ij}, r, \beta$ in (9). Note that conditioning on $z_{ij}$ is equivalent to conditioning on $\mathbf{z}$ as long as $r$ and $\beta$ are known. Then, we can estimate the first player's coverage probability by $\sum_{i=1}^{500}E(I_{i1}\vert z_{i1}, r, \beta)/500$, which is also unbiased but more accurate than the previous simple estimator.

\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's abililty to handle this surgery, it will be useful for your decision in choosing a hospital.
\\

@@

Intuitively, the result of multilevel modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose sample means ($z_{j} / n_{j}$) are 0 and 0.001 and caseloads ($n_{j}$) are 1 and 1000 each. Do you believe that the former hospital, whose observed death rate is 0, is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge its true death rate per exposure with just one caseload. Though somewhat extreme, this is what happened to the two left-most hostpitals in the first plot and this is why hierarchical modeling is a reasonable choice for this data set.
\\

@@

Also it reveals that the 23rd hospital, whose estimated true rate was the smallest, has also the smallest upper bound. If you are a risk-avoider, this hospital will attract you most strongly. And if you already chose this 23rd hospital compared to the 8th from the shrinkage plot, your decision might become stronger at this point, excluding the 8th hospital with more certainty. 
\\

@@

It estimated coverage probabilities by simple unbiased estimates (a blue dashed line) and Rao-Blackwellized unbiased estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.950 from the Rao-Blackwellized estimate and 0.951 from the simple unbiased estimate. Considering this result, we finally conclude that the interval estimates from the suggested multilevel modeling on this particular data set has nice repeated sampling property.
\\

@@

One more thing to emphasize is that observed sample means are all within the interval estimates. And these interval estimates also catched the batting averages over the ramaining season successfully, although the result was left out. Then, how much can we trust these interval estimates? The following method check will answer it.

>>>>

Considering the type of data, it makes sense to assume the number of deaths in each hospital has independent Poisson distribution given the unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where $n_{j}$ can be interpreted as an exposure (not necessarily an integer) and $\lambda_{j}$ as a true death rate per exposure. This data set is also included in the package and can be called by `\code{R> data(hospital)}' on \proglang{R}.

@@

\section[Multilevel Structure]{Multilevel Structure}
Our multilevel model, also called a conditionally independent hierarchical model (Kass and Steffey, 1989), is a very powerful tool for exploring the hierarchical structure in data. For example, we can think about a district-level hierarchy (bigger population) for 8 schools, the state-level hierarchy for 31 hospitals, and the position-level hierarchy for 18 baseball players. \code{gbp}, one of functions in \pkg{Rgbp}, fits such a hierarchical model whose first-level hierarchy has a distribution of observed data and second-level (bigger population hierarchy) has a conjugate prior distribution on the first-level parameter. Users can determine one of three types of multilevel models, such as Normal-Normal, Poisson-Gamma, and Binomial-Beta, based on their data sets. 
\\

@@

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only data. Taking the Normal model as an example,  $E(\mu_{j}\vert \textbf{y})=E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Adam's law and $Var(\mu_{j}\vert \textbf{y})=E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Eve's law. Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are functions of shrinkage (see (3)) and Adam's and Eve's laws are taking conditional expectation and variance over them given data, $\textbf{y}$. Since we have already estimated $E(B_{j}\vert\textbf{y})$ and $Var(B_{j}\vert\textbf{y})$ via ADM, finally we can estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$.

@@

Once we are sure that the hierarchical modeling can be appropriate for our data, the following question will be about the validity of interval estimates, the final product of this multilevel modeling. ``Does the 95\% (can be specified differently) confidence interval obtained via this Bayesian model-fitting process achieve 95\% confidence level  for any true parameter values?'' Our answer is ``yes'' and \pkg{Rgbp} has a function to check this point. From now on, all the explanations will be based on the Binomial model because the idea is the same and can be easily applied to the other two models.

@@

\subsection{Rao-Blackwellized Unbiased estimates of Coverage Probabilities}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking conditional expectation given a sufficient statistic. Based on this idea, we define $E(I_{ij}\vert z_{ij}, r, \beta$), where $r$ and $\beta$ were given at first (see 5.1) and $z_{ij}$ is a sufficient statistic. This expectation is the same as Pr$(\hat{p}_{ij, low}\le p_{ij} \le\hat{p}_{ij, upp}\vert z_{ij}, r, \beta)$, where ($\hat{p}_{ij, low}$, $\hat{p}_{ij, upp}$) is $j$-Ch player's interval estimates on the $i$-th data set. We can calculate this probability  because we know the distribution of $p_{ij} \vert z_{ij}, r, \beta$ in (9). Note that conditioning on $z_{ij}$ is equivalent to conditioning on $\mathbf{z}$ as long as $r$ and $\beta$ are known. Then, we can estimate the first player's coverage probability by $\sum_{i=1}^{500}E(I_{i1}\vert z_{i1}, r, \beta)/500$, which is also unbiased but more accurate than the previous simple estimator.

\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's ability to handle this surgery, it will be useful for your decision in choosing a hospital.
\\

@@

Intuitively, the result of multilevel modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose sample means ($z_{j} / n_{j}$) are 0 and 0.001 and caseloads ($n_{j}$) are 1 and 1000 each. Do you believe that the former hospital, whose observed death rate is 0, is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge its true death rate per exposure with just one caseload. Though somewhat extreme, this is what happened to the two left-most hospitals in the first plot and this is why hierarchical modeling is a reasonable choice for this data set.
\\

@@

Also it reveals that the 23rd hospital, whose estimated true rate was the smallest, has also the smallest upper bound. If you are a risk-adverse, this hospital will attract you most strongly. And if you already chose this 23rd hospital compared to the 8th from the shrinkage plot, your decision might become stronger at this point, excluding the 8th hospital with more certainty. 
\\

@@

It estimated coverage probabilities by simple unbiased estimates (a blue dashed line) and Rao-Blackwellized unbiased estimates (a red line). Both lines are indistinguishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.950 from the Rao-Blackwellized estimate and 0.951 from the simple unbiased estimate. Considering this result, we finally conclude that the interval estimates from the suggested multilevel modeling on this particular data set has nice repeated sampling property.
\\

@@

One more thing to emphasize is that observed sample means are all within the interval estimates. And these interval estimates also covered the batting averages over the remaining season successfully, although the result was left out. Then, how much can we trust these interval estimates? The following method check will answer it.
